Final Report

### 2022-05-10 Final Report.Rmd
graph explanations

North America: U.S, Canada
South America: Brazil, Chile
Africa: Nigeria, Malawi
Europe: Germany, Serbia
Asia: Pakistan, Sri Lanka
Oceania: Australia, New Zealand

@@

Lastly, we will join the population and and deaths with its respected country.

@@

```{r}
joined_all <- right_join(deaths_df, world_pop, by=c('country' = 'Country.Name', 'year' = 'Year'))
head(joined_all)

north_america <- joined_all %>% filter(country %in% c("United States", "Canada"))
head(na.omit(north_america))

south_america <- joined_all %>% filter(country %in% c("Brazil", "Chile"))
head(na.omit(south_america))

africa <- joined_all %>% filter(country %in% c("Nigeria", "Malawi"))
head(na.omit(africa))

europe <- joined_all %>% filter(country %in% c("Germany", "Serbia"))
head(na.omit(europe))

asia <- joined_all %>% filter(country %in% c("Pakistan", "Sri Lanka"))
head(na.omit(asia))

oceania <- joined_all %>% filter(country %in% c("Australia", "New Zealand"))
head(na.omit(oceania))

@@

ggplot(data=joined_low, aes(x=year, y=Count, color=country, size = total_deaths/Count)) +
  geom_line() +
  labs(title="Percentage of Death Count in Low Population Countries", subtitle="Bar width is equal to percentage of Deaths")+
  xlab("Year")+
  ylab("Population")
```
These graphs are of the same information but we have added the percentage of air-pollution related deaths as the width of the line to demonstrate visually if deaths increased or decreased over time. It is easier to see in the high-populated countries but air-pollution related deaths do decrease over time.

@@

```
We wanted to take a closer look at the death count and see which country has the highest average death count. In the tables we made you can see that Pakistan had the highest average death count at 144.33 for the high populated countries. Malawi had the highest average death count at 147.77 for the low populated countries, which is higher than Pakistan.

@@

@@

kable(high_poll)

#Low Population Pollutant Averages 
low_poll <- deaths_df %>% 
  group_by(country) %>% 
  filter(country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand')) %>% 
  select(country, indoor_deaths, outdoor_deaths, ozone_deaths) %>% 
  summarize(avg_indoor = mean(indoor_deaths), avg_outdoor = mean(outdoor_deaths), avg_ozone = mean(ozone_deaths))

@@

@@

#Low Indoor Air Pollution
l_indoor<- ggplot(low_poll, aes(x=country, y = avg_indoor, cex=5)) + 
  geom_point() +
  ggtitle("Indoor Air Pollution Deaths in Low Population Countries")+
  xlab("Country")+
  ylab("Average Deaths")
ggplotly(l_indoor)

# High Outdoor Air Pollution
h_outdoor <- ggplot(high_poll, aes(x=country, y = avg_outdoor, color = avg_ozone, cex=5)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "violet", aesthetics = "colour") +
  geom_point() +
  labs(title = "Outdoor Air Pollution Deaths in High Population Countries") +
  xlab("Country") + 
  ylab("Average Deaths")
ggplotly(h_outdoor)

@@

>>>>

  + North America: U.S, Canada
  + South America: Brazil, Chile
  + Africa: Nigeria, Malawi
  + Europe: Germany, Serbia
  + Asia: Pakistan, Sri Lanka
  + Oceania: Australia, New Zealand

@@

Lastly, we will join the population and and deaths with its respected country. The last column displays the 'Count' which represents the population of that country at that time period.

@@

```{r}
joined_all <- right_join(deaths_df, world_pop, by=c('country' = 'Country.Name', 'year' = 'Year'))
head(joined_all)
```

North America
```{r}
north_america <- joined_all %>% filter(country %in% c("United States", "Canada"))
head(na.omit(north_america))
```

South America 
```{r}
south_america <- joined_all %>% filter(country %in% c("Brazil", "Chile"))
head(na.omit(south_america))
```

Africa
```{r}
africa <- joined_all %>% filter(country %in% c("Nigeria", "Malawi"))
head(na.omit(africa))
```

Europe
```{r}
europe <- joined_all %>% filter(country %in% c("Germany", "Serbia"))
head(na.omit(europe))
```

Asia
```{r}
asia <- joined_all %>% filter(country %in% c("Pakistan", "Sri Lanka"))
head(na.omit(asia))
```

Oceania
```{r}
oceania <- joined_all %>% filter(country %in% c("Australia", "New Zealand"))
head(na.omit(oceania))

@@

ggplot(data=joined_low, aes(x=year, y=Count, color=country, size = total_deaths/Count)) +
  geom_line() +
  labs(title="Percentage of Death Count in Low Population Countries", subtitle="Bar width is equal to percentage of Deaths")+
  xlab("Year")+
  ylab("Population")
```

These graphs are of the same information but we have added the percentage of air-pollution related deaths as the width of the line to demonstrate visually if deaths increased or decreased over time. It is easier to see in the high-populated countries but air-pollution related deaths do decrease over time.

@@

We wanted to take a closer look at the death count and see which country has the highest average death count. In the tables we made you can see that Pakistan had the highest average death count at 144.33 for the high populated countries. Malawi had the highest average death count at 147.77 for the low populated countries, which is higher than Pakistan.

@@

Looking between indoor, outdoor, and ozone pollution deaths, we can see which pollutant-type had the greatest death count.

@@

kable(high_poll)
```
The average indoor-death count is higher than other pollutant types, especially in Pakistan and Nigeria. In all three categories, Pakistan had the highest death count.

```{r}
#Low Population Pollutant Averages 
low_poll <- deaths_df %>% 
  group_by(country) %>% 
  filter(country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'New Zealand')) %>% 
  select(country, indoor_deaths, outdoor_deaths, ozone_deaths) %>% 
  summarize(avg_indoor = mean(indoor_deaths), avg_outdoor = mean(outdoor_deaths), avg_ozone = mean(ozone_deaths))

@@

In the low populated countries, Malawi has a higher average indoor death than that of any of our other selected countries. It is also the highest in all three categories, as well.

@@

#Low Indoor Air Pollution
l_indoor<- ggplot(low_poll, aes(x=country, y = avg_indoor, cex=5)) + 
  geom_point() +
  ggtitle("Indoor Air Pollution Deaths in Low Population Countries")+
  xlab("Country")+
  ylab("Average Deaths")
ggplotly(l_indoor)
```
In these graphs, it is much easier to see the discrepancies in the indoor deaths of the countries. Malawi, Pakistan, and Nigeria have dots located high on the graph indicating high death counts.

```{r}
# High Outdoor Air Pollution
h_outdoor <- ggplot(high_poll, aes(x=country, y = avg_outdoor, color = avg_ozone, cex=5)) +
  scale_color_gradient2(low = "light pink", mid = "pink", high = "violet", aesthetics = "colour") +
  geom_point() +
  labs(title = "Outdoor Air Pollution Deaths in High Population Countries") +
  xlab("Country") + 
  ylab("Average Deaths")
ggplotly(h_outdoor)

@@

In these graphs, the outdoor pollution deaths are displayed on the y-axis with the amount of ozone-related deaths indicated by the color gradient. The dots highest-up on the graph have high outdoor-related deaths, and the darker the color of pink indicates a high ozone-related death.

For our high-populated countries, Pakistan had high outdoor-deaths and ozone-deaths while the other high-populated countries were lower in ozone-deaths. In the low-populated countries, Serbia had the highest outdoor and ozone deaths, but Malawi in particular had the greatest ozone deaths.
